
### IRIS DATA

```{r,echo=FALSE,message=FALSE,warning=FALSE,results='asis'}
library(PPforest)
library(PPtree)
library(reshape)
library(plyr)
library(ggplot2)
library(randomForest)
library(lineprof)
library(PPtree)
library(rpart)
library(randomForest)


```

```{r data,echo=FALSE,cache=TRUE,message=FALSE,warning=FALSE,results='asis'}
#IRIS data example

data.iris <- iris[,5:1]
#training <- train_fn(iris[,5],.9)
#test <- as.vector(1:length(iris[,5]))[!(1:length(iris[,5])%in%(training))]


result.boot.st <- mlply(data.frame(nt=c(5,10,50,100,500)), function(nt) {
  bootstrap_pp(data=data.iris,scale=TRUE,size.p=.9,training=NULL,test=NULL,strata=TRUE,ntree=nt,index="LDA") 
   }  
)  


result.bagg.st <- llply(result.boot.st[1:5],function(x)  bagging_pp(data.iris,scale=TRUE, strata=TRUE,output,training=NULL,test=NULL))

error.st <- sapply(result.bagg.st,function(x) x[[1]][1] )


bagging_pp(data.iris,scale=TRUE, strata=TRUE,output,training,test)
#proximity matrix
A <- matrix(0,ncol=150,nrow=150)
A[lower.tri(A,diag=T)]<-b.pp[[9]][,3]
A[upper.tri(A,diag=T)]
dA <- as.dist(1-A)
clA <- hclust(dA)
cl.h <- cutree(clA, k = 3)
qplot(y=Sepal.Width, x=Sepal.Length, data=iris,color=as.factor(cl.h))


km.A <- kmeans(A,centers=3)
qplot(y=Species, x=km.A$cluster,data=iris, group=km.A$cluster,geom="jitter" )
table(iris$Species, km.A$cluster)
qplot(y=Sepal.Width, x=Sepal.Length, data=iris,color=as.factor(km.A$cluster))
qplot(y=Petal.Length, x=Petal.Width, data=iris,color=as.factor(km.A$cluster))
qplot(y=Sepal.Length, x=Petal.Width, data=iris,color=as.factor(km.A$cluster))


```

```{r table.res,echo=FALSE,dependson='data',results='asis'}
library(xtable)
aux <- c(5,10,50,100,500)

aaa <- xtable(cbind(aux,error.st),digits=5,caption="OOB error for PP.bagging with bootstrap samples without strata and with strata")
###
print(aaa,caption.placement="top",include.rownames=TRUE,include.colnames=TRUE)

```

Using PPtree with LDA the error is 0

```{r pp.iris,dependson='data',echo=TRUE,message=FALSE,warning=FALSE}
library(PPtree)
Tree.result <- PP.Tree("LDA",iris[training,5],iris[training,1:4])
test <- iris[-training,]
res<-PP.classify(test[,1:4],test[,5],Tree.result,1)
print(res[[1]]/length(res[[2]]))
```

Ussing the same data we run a random forest and the oob error is around 5%.

```{r rf.iris,depenson="data'",echo=FALSE,message=FALSE,warning=FALSE}
library(randomForest)
rf.iris<-randomForest(Species ~ ., data=iris[training,], importance=TRUE,
                     proximity=TRUE,strata=Specie)
print(rf.iris)
```



### OLIVE DATA

```{r olive,cache=TRUE,echo=FALSE,results='asis',message=FALSE,warning=FALSE}
setwd("/Users/nataliadasilva/Documents/PP.forest/PP.Forest/data")
load("olive.rda")
library(ggplot2)

d.olive2 <- subset(d.olive,Region%in%c(1,2))[,-2]
d.olive2$Region <- factor(d.olive2$Region)

training <- train_fn(d.olive2[,1],.9)

boot<-bootstrap_pp(data=d.olive2,training=training,ntree=50,index="LDA") 
 b.pp <- bagging_pp(d.olive2,scale=TRUE, strata=TRUE,boot,training)

  

result.boot2.st <- mlply(data.frame(nt=c(5,10,50,100,500)), function(nt) {
  bootstrap_pp(data=d.olive2,scale=TRUE,size.p=.9,training=training,strata=TRUE,ntree=nt,index="LDA") 
   }  
)  

 

result.bagg2.st <- llply(result.boot2.st[1:5],function(x) bagging_pp(data=d.olive2,scale=TRUE,strata=TRUE,boot=x,training=training))

error.st <- sapply(result.bagg2.st,function(x) x[[1]][1] )

```


```{r table.olive,echo=FALSE,dependson='olive',results='asis',message=FALSE,warning=FALSE}
library(xtable)
aux <- c(5,10,50,100,500)

aaa<-xtable(cbind(aux,error.st),digits=5,caption="OOB error for PP.bagging with bootstrap samples without strata and with strata")
###
print(aaa,caption.placement="top",include.rownames=TRUE,include.colnames=TRUE)

```


```{r pp.olive,dependson='olive',cache=TRUE,echo=FALSE}
library(PPtree)
Tree.result <-  LDA.Tree(d.olive2[training,1],d.olive2[training,2:9])



test <- d.olive2[-training,]
res<-PP.classify(d.olive2[training,2:9],d.olive2[training,1],Tree.result,1)

print(res[[1]]/length(res[[2]]))

```

Using random forest the OOb error us 0\%

```{r rf.olive,dependson='olive',cache=TRUE,echo=FALSE}
library(randomForest)

rf <- randomForest(Region ~ ., data=d.olive2[training,], importance=TRUE,strata=Region,
             proximity=TRUE)

print(rf )

```


Basic Output of PPforest for ntree=50

```{r olive2,depenson='olive',cache=TRUE,echo=FALSE}

boot<-bootstrap_pp(data=d.olive2,training=training,ntree=50,index="LDA") 
 b.pp <- bagging_pp(d.olive2,scale=TRUE, strata=TRUE,boot,training)


```




```{r data.cv,cache=TRUE,echo=FALSE}
setwd("./data")
crab.aux <- read.csv("australian-crabs.csv",header=T)[,-3]
crab <-data.frame(Type=with(crab.aux,paste(species,sex,sep="")),crab.aux[,-c(1,2)])
leukemia <- read.csv( "Leuk-tot.csv",header=T)
lymphoma <- read.csv("Lymp-tot.csv",header=T)
NCI60 <- read.csv("NCI60.csv",header=T)
wine <- read.csv( "wine.csv", header=T)
glass <- read.csv("glass.csv",header=T)
fishcatch <- read.csv("fishcatch.csv",header=T)[,-1]
names(fishcatch)[1]<- "Type"
image <- read.csv("image.csv",header=T  )
parkinson <- read.csv("parkinsons.csv",header=T)

#temp = list.files(pattern="*.csv")
#myfiles = lapply(temp, read.delim)
```

```{r datapp,cache=TRUE,echo=FALSE,dependson="data.cv"}


training <- train_fn(leukemia [,1],.9)

result.boot2.st <- mlply(data.frame(nt=c(5,10,50,100,500)), function(nt) {
  bootstrap_pp(data=leukemia,scale=TRUE,size.p=.9,training=training,strata=TRUE,ntree=nt,index="LDA") 
   }  
)  

 result.bagg2.st <- llply(result.boot2.st,function(x) bagging_pp(data=leukemia,scale=TRUE,strata=TRUE,boot=x,training=training))

error.st <- sapply(result.bagg2.st,function(x) x[[1]][1] )

aux <- c(5,10,50,100,500)

aaa<-xtable(cbind(aux,error.st),digits=5,caption="OOB error for PP.bagging with bootstrap samples without strata and with strata")
###
print(aaa,caption.placement="top",include.rownames=TRUE,include.colnames=TRUE)

```

 


#Corss validation

```{r cv,dependson="data.cv",cache=TRUE,echo=FALSE}


#PPtree
lda.err <- function(tr,te){
ppt.lda <-LDA.Tree(tr[,1],tr[,-1])

ppt.pred.tr <-PP.classify(tr[,-1], tr[,1],ppt.lda , Rule=1)

ppt.pred.te <-PP.classify(te[,-1], te[,1],ppt.lda , Rule=1)

 err.tr <- ppt.pred.tr[[1]]/length(ppt.pred.tr[[2]])

 err.te <- ppt.pred.te[[1]]/length(ppt.pred.te[[2]])
data.frame(err.tr,err.te)

}

#PPtree
pda.err <- function(tr,te){
ppt.pda <-PDA.Tree(tr[,1],tr[,-1],lambda=.14)

ppt.pred.tr <-PP.classify(tr[,-1], tr[,1],ppt.pda , Rule=1)

ppt.pred.te <-PP.classify(te[,-1], te[,1],ppt.pda , Rule=1)

 err.tr <- ppt.pred.tr[[1]]/length(ppt.pred.tr[[2]])

 err.te <- ppt.pred.te[[1]]/length(ppt.pred.te[[2]])
data.frame(err.tr,err.te)

}


#CART
cart.err <- function(tr,te){
cart<- rpart(as.factor(Type)~.,data=tr)
fit.preds.tr <- predict(cart,newdata=tr,type="class")
fit.preds.te <- predict(cart,newdata=te,type="class")
fit.table.tr <- table(tr[,1],fit.preds.tr)
fit.table.te <- table(te[,1],fit.preds.te)
err.tr <- (dim(tr)[1]-sum(diag(fit.table.tr)))/dim(tr)[1]
err.te <- (dim(te)[1]-sum(diag(fit.table.te)))/dim(te)[1]
data.frame(err.tr,err.te)
}

#RF
rf.err <- function(tr,te){
rf <- randomForest(as.factor(Type) ~ ., data=tr)
 err.tr <- (dim(tr)[1]-sum(diag(rf$confusion[,-dim(rf$confusion)[2]])))/dim(tr)[1]
 fit.preds.te <- predict(rf,newdata=te,type="class")
fit.table.te <- table(te[,1],fit.preds.te)
err.te <- (dim(te)[1]-sum(diag(fit.table.te)))/dim(te)[1]
data.frame(err.tr,err.te)
}

#PP.RF

pprf.err <- function(data,tr,te,tr.index,te.index,lambda=0.14,strata=TRUE,index="LDA"){
bo.pp <- bootstrap_pp(data=data,scale=TRUE,size.p=.9,training=tr.index,strata=strata,ntree=5,index=index,lambda=0.14)      
ba.pp <- bagging_pp(data,scale=TRUE,strata=strata,bo.pp,tr=tr.index,te=te.index)
data.frame(ba.pp[[2]],ba.pp[[3]])
}


cv.meth <- function(data) {

rdply(2 , function(){
  tr.index <- train_fn(data[,1],2/3)
  te.index <- as.vector(1:length(data[,1]))[!(1:length(data[,1])%in%(tr.index))]
 training<- data[tr.index,]
test <- data[-tr.index,] 
data.frame(lda=lda.err(training,test),cart=cart.err(training,test),rf=rf.err(training,test),pprf=pprf.err(data,training,test,tr.index,te.index)) 
})
}

```

\end{document}

```{r}
data<-leukemia

 tr.index <- train_fn(data[,1],2/3)
  te.index <- as.vector(1:length(data[,1]))[!(1:length(data[,1])%in%(tr.index))]
 training<- data[tr.index,]
test <- data[-tr.index,] 
ppt.lda <-LDA.Tree(training[,1],training[,-1])


PDA.Tree(training[,1],training[,-1],lambda=.14)

table(training[,36])
#LDA in some cases the sample  Error in lda.default(x, grouping, ...) : 
 # variable 27 appears to be constant within groups
#PDA does not work with any lambda, why?
#How you get the cv for leukemia data if you have this problem because you use lda, pda

data<-NCI60
training<-train_fn(data[,1],.9)
#part of bootstrap_pp code
data[,-1] <- scale(data[,-1])
names(data)[1] <-"class"

 n <- length(training)
    class.id <- data.frame(id=1:n,class=data[training,"class"])
    dat.train <- data[training,]

index.boot <- unlist(plyr::dlply(class.id, plyr::.(class), function(x) sort(sample(x$id, replace=TRUE)) ))
    names(index.boot) <- NULL 
    pp.tree <-PPtree_split(PPmethod=index, size.p=.9,i.data=dat.train[index.boot,-1], i.class=dat.train[index.boot,1]) 
    pp.tree <-PP.Tree(PPmethod=index, size.p=.9,i.data=dat.train[index.boot,-1], i.class=dat.train[index.boot,1]) 

pp.tree

#I have a problem in the way I select the variables or how I change it in pptree function for some data work but not for all
```


Profilling
``{r}
data1 <- iris[,5:1]

 training <- train_fn(iris[,5],.9)
 test <- as.vector(1:length(iris[,5]))[!(1:length(iris[,5])%in%(training))]
 output <- bootstrap_pp(data1,scale=TRUE,size.p=.9,training=NULL,strata=TRUE,ntree=50,index="LDA")
 
 bag.t <- system.time(bagging_pp(data1,scale=TRUE, strata=TRUE,output,training,test))
boo.t <- system.time(bootstrap_pp(data1,scale=TRUE,size.p=.9,training=NULL,strata=TRUE,ntree=50,index="LDA"))

pptree.sp.t <- system.time(PPtree_split("LDA", data1[training,1], data1[training,2:5],size.p=0.9))


ptm <- proc.time()
b1<-bagging_pp(data1,scale=TRUE, strata=TRUE,output,training,test)
proc.time() - ptm

ptm <- proc.time()
b2<-bootstrap_pp(data1,scale=TRUE,size.p=.9,training=NULL,strata=TRUE,ntree=50,index="LDA")
proc.time() - ptm

library(profr)

bl1 <- profr(
  b1<-bagging_pp(data1,scale=TRUE, strata=TRUE,output,training,test)
)
head(bl1, 10)

plot(bl1)

bl2 <- profr(
  b2<-bootstrap_pp(data1,scale=TRUE,size.p=.9,training=NULL,strata=TRUE,ntree=50,index="LDA")
)
head(bl2, 20)
plot(bl2)

ldat <- profr(
  lda<-LDA.Tree(iris[,5],iris[,-5])
)
head(ldat, 20)
plot(ldat)

```



```